Agile Monitoring of Evolving Web Services

### 2018-04-03 evolution.tex
cleaning up

\begin{figure*}[h!]
  \centering
  \subfloat[overview of the tracked endpoints]{\includegraphics[width=1.5\columnwidth]{overview}\label{fig:overview}}
  \caption{Some of the available views\label{fig:views}}
\end{figure*}

  \newpage
  \section{Evolution monitoring }

@@

  If the \tool can automatically detect the current version of the project by reading the .git configuration as soon as the API is started and can then group measurements by version\footnote{Alternatively, the maintainer can add version identifiers manually for the web application through a configuration file if the system does not use git.}. \Fref{fig:tee} is a zoomed-in version of such a view for \epTranslations with versions increasing from top to bottom

@@

\begin{figure*}[!ht]
  \centering
  \includegraphics[width=0.4\linewidth]{time_per_user}
  \caption{The \epFeedItems shows a very high variability across users}
  \label{fig:tpu}
\end{figure*}

>>>>

  \section{Evolution monitoring }

@@

  If the \tool can automatically detect the current version of the project by reading the .git configuration as soon as the API is started and can then group measurements by version\footnote{Alternatively, the maintainer can add version identifiers manually for the web application through a configuration file if the system does not use git.}.

  \subsubsection*{Evolving Utilization}

  \Fref{fig:mv-util} shows the \perspective{Multi-Version Utilization} perspective. 
  It shows the utilization of all the tracked endpoints across versions.

    \begin{figure}[h!]
      \centering
      \includegraphics[width=\linewidth]{utilization-evolution}
      \caption{The Evolution of Endpoint Utilization Across System Versions}
      \label{fig:mv-util}
    \end{figure}

    \subsubsection*{Evolving Performance}

    \Fref{fig:tee} is a zoomed-in version of such a view for \epTranslations with versions increasing from top to bottom

@@

\subsubsection*{Evolving Groups}
The limitation of the previous view is that it does not present the information also on a per version basis. To address this, a different visual perspective entitled \perspective{Multi-Version per-User Performance} can be defined. Figure \ref{fig:tuv} presents such a perspective by mapping the average execution time for a given user (lines) and given version (columns) on the area of the corresponding circle.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.7\linewidth]{time_per_user_per_version}
  \caption{This perspective shows that the evolution of response times for individual users (horizontal lines) across versions (the x-axis)}
  \label{fig:tuv}
\end{figure}

The colors represent users. The figure shows average performance varying  across users and versions with no clear trend: this is probably because varying user workload (i.e. number of sources to which the user is registered) is the reason for the variation in response times. \ins{one can see that for user 1 performance degrades over the }

### 2018-04-03 grouping.tex
cleaning up

\newpage
\section{Grouping Requests}
\label{sec:user}

For service endpoints which run computations in real time, the maintainer of a system might want to understand the endpoint performance on a per-user basis, especially for situations where the system response time is a function of some individual user load\footnote{E.g. in GMail some users have two emails while other have twenty thousand and this induces different response times for different users}.

\todo{remove the configuration part and rephrase the following, the discussion has been partially subsumed by Section~\ref{sec:tool}.}

@@

%\begin{lstlisting}[float,caption=Simply define a custom app-specific function for user retrieval and pass it to the \tool to group information by user,style=custompython]
\begin{lstlisting}[style=custompython]  
# app specific way of extracting the user
# from a flask request object    
def get_user_id():
sid = int(flask.request.args['session'])
session = User.find_for_session(sid)
return user_id

# attaching the get_user_id function
dashboard.config.get_group_by = get_user_id

@@

The limitation of the previous view is that it does not present the information also on a per version basis. To address this, a different visual perspective entitled \perspective{Multi-Version per-User Performance} can be defined. Figure \ref{fig:tuv} presents such a perspective by mapping the average execution time for a given user (lines) and given version (columns) on the area of the corresponding circle.

The colors represent users. The figure shows average performance varying  across users and versions with no clear trend: this is probably because varying user workload (i.e. number of sources to which the user is registered) is the reason for the variation in response times. \ins{one can see that for user 1 performance degrades over the }

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.7\linewidth]{time_per_user_per_version}
  \caption{This perspective shows that the evolution of response times for individual users (horizontal lines) across versions (the x-axis)}
  \label{fig:tuv}
\end{figure}

>>>>

\section{Grouping Requests}
\label{sec:user}

For service endpoints which run computations in real time, the maintainer of a system might want to understand the endpoint performance on a per-user basis, especially for situations where the system response time is a function of some individual user load\footnote{E.g. in GMail some users have two emails while other have twenty thousand and this induces different response times for different users}.
\todo{remove the configuration part and rephrase the following, the discussion has been partially subsumed by Section~\ref{sec:tool}.}

@@

\begin{lstlisting}[style=custompython]

# attaching a group by function 
dashboard.config.group_by = 'User',
	lambda: Session.find(flask.request).userid

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=.8\linewidth]{time_per_user}
  \caption{The \epFeedItems shows a very high variability across users}
  \label{fig:tpu}
\end{figure}

### 2018-04-03 performance.tex
cleaning up

@@

  \newpage
  \subsection{Performance Insight: Automated outlier detection and monitoring}

@@

The bottom panel shows the stack trace. 
In this particular case, it is revealing for the developer to learn that at the time of the stack trace snapshot, the code was in the microsoft\_translator.py: indeed, the system uses as back-end multiple translators, and it has been observed that many of the outliers happen to be waiting in the microsoft translator.

@@

\begin{figure*}[h!]
  \centering
  \includegraphics[width=0.9\linewidth]{utilization-evolution}
  \caption{The Evolution of Endpoint Utilization Across System Versions}
  \label{fig:tee}
\end{figure*}

>>>>

Knowing the performance of one’s API is critical for the quality of a service API. Flask Dashboard introduces a series of perspectives on performance, which focus on the response times of the various endpoints.

\ins{HERE? ELSEWHERE? Actually, probably better for performance. The importance of the last day and last seven days became clear only after having used the dashboard for several months. }

@@

  \newpage
  \subsection{Automated Outlier Detection and Monitoring}

@@

The bottom panel shows the stack trace. 
In this particular case, it is revealing for the developer to learn that at the time of the stack trace snapshot, the code was in the google\_translator.py: indeed, the system uses as back-end multiple translators, and it has been observed that many of the outliers happen to be waiting in the google translator.

@@

### 2018-04-03 testing.tex
cleaning up

  \newpage
  \section{Preemptive monitoring}

@@

@@

    \begin{figure}[h!]
        \centering
        \includegraphics[width=\columnwidth]{travis_builds}
        \caption{Response times in 5 consequent Travis builds in the \zee case study}
        \label{fig:builds}
      \end{figure}

@@

\begin{figure}[h!]
  \centering
  \subfloat[The reported response time (in ms) per deployed version in the observation period]{\includegraphics[width=.5\columnwidth]{response_per_version_trunced_trend}}
  \quad
  \subfloat[The measured response time (in ms) using integration with Travis]{\includegraphics[width=.46\columnwidth]{travis_builds_no_outliers_trend}}
  \caption{Comparison of the response times per endpoint: actual production system versus preemptive monitoring data}
  \label{fig:preemptive}
\end{figure}

>>>>

  \begin{figure*}[h!]
  \centering
  \subfloat[The reported response time (in ms) per deployed version in the observation period]{\includegraphics[width=.6\columnwidth]{response_per_version_trunced_trend}}
  \quad
  \subfloat[The measured response time (in ms) using integration with Travis]{\includegraphics[width=.56\columnwidth]{travis_builds_no_outliers_trend}}
  \caption{Comparison of the response times per endpoint: actual production system versus preemptive monitoring data}
  \label{fig:preemptive}
\end{figure*}

  \section{Preemptive monitoring}

@@

      \begin{figure}[h!]
        \centering
        \includegraphics[width=0.7\columnwidth]{travis_builds}
        \caption{Response times in 5 consequent Travis builds in the \zee case study}
        \label{fig:builds}
      \end{figure}

@@

@@

### 2018-04-03 utilization.tex
cleaning up

  The most basic possible view shows the cummulative information about all the endpoints and the number of calls to that endpoint (over the lifetime of its tracking as well as in the current day and the last seven days). HERE? ELSEWHERE? Actually, probably better for performance. The importance of the last day and last seven days became clear only after having used the dashboard for several months. 
}

@@

This kind of information is already very useful, and can provide the API maintainer with insight into the evolution of their system. By looking at \Fref{fig:basicest}, which presents the top 7 endpoints (for lack of space) one can already see several patterns: 
- the two most used endpoints stand for the two main activities in the system (doing exercises and asking for translations while reading personalized materials )
- \epUserActivity has been disproportionately been used in the last day and last week
- \epFeedItems has not been used in the last 7 days

@@

@@

Figure \ref{fig:aeu} shows a first perspective on endpoint utilization that \tool provides: a stacked bar chart of the number of hits to various endpoints grouped by day\footnote{Endpoint colors are the same in different views}. Figure~\ref{fig:aeu} in particular shows that at its peak the API has about 12.000 hits per day.

@@

% \begin{enumerate}

%   \item {\color{myred} \epTranslations} is an indicator of the amount of foreign language reading the users are doing, and

%   \item {\color{mygreen} \epOutcome} is an indicator of the amount of foreign vocabulary practice the users are doing.

% \end{enumerate}

@@

A second type of utilization question that the \tool can answer automatically regards {\em cyclic patterns of usage per hour of day} by means of a heatmap, as in \Fref{fig:dp}.

% \mltp{ can we add vertical lines that highlight the beginning of a new week (e.g. before Sunday): }
% Patrick: adding separators in the graph is unfortunately not supported by the library.

>>>>

  The most basic possible view shows the cummulative information about all the endpoints and the number of calls to that endpoint over the lifetime of its tracking as well as in the current day and the last seven days.  
}

@@

This kind of information is already very useful, and can provide the API maintainer with insight into the evolution of their system. By looking at \Fref{fig:basicest}, which presents the top 7 endpoints (for lack of space) one can already see several patterns:

\begin{itemize}

  \item {\bf Frequent Patterns of System Usage}. The two most used endpoints stand for the two main activities in the system:

  \begin{itemize}

    \item \epTranslations is an indicator of the amount of foreign language reading the users are doing

    \epOutcome is an indicator of the amount of foreign vocabulary practice the users are doing.

  \end{itemize}

  \item {\bf Sudden Increase of Endpoint Usage}. One endpoint (\epUserActivity) has been disproportionately been used in the last day and last week; much more than before.

  \item {\bf Possibly Discontinued Endpoint Usage}. One endpoint (\epFeedItems) has not been used in the last 7 days

\end{itemize}

\lesson{Although very useful, it was only after several months of using the system that the client requested the extra columsn for one day and seven days.}

@@

The table view is limited, and the one day and one week periods are conveniently but arbitrarily selected. For a more detailed evolution of utilization, Figure \ref{fig:aeu} shows the \perspective{Daily Utilization} perspective on endpoint utilization that \tool provides: a stacked bar chart of the number of hits to various endpoints grouped by day\footnote{Endpoint colors are the same in different views}. Figure~\ref{fig:aeu} in particular shows a peak utilization, a day when the API had more than 12.000 hits.

@@

@@

@@

Another utilization perspective is the \perspective{Hourly Utilization} in which  the \tool can highlight {\em cyclic patterns of usage per hour of day} by means of a heatmap, as in \Fref{fig:dp}.
